### 2015-02-02 paper/conclusion.tex
saving x_true,x_est,f with output; adding paper template

%!TEX root = main.tex

\section{Conclusion}
The conclusion goes here.