### 2015-02-02 paper/intro.tex
+ bibliography

%!TEX root = main.tex

\section{Introduction}
In this work, we study the noise sensitivity of the state-of-the-art for route flow estimation for mobility cyber-physical systems (CPS). Route flow estimation on road networks is a critical problem for mobility CPS, as any prediction and control applications require accurate and reliable estimates of the state of the road network. Previous work has demonstrated the promise of using cellular towers as sensor inputs to a convex optimization formulation of the problem, along with a projected first-order method that scales well to large city-scale networks \cite{Wu2015}. However, the work uses entirely synthetic data which assumes that vehicles are connected to the closest cell tower throughout their trajectories (i.e. voronoi partitions). In