### 2015-02-02 paper/experiments.tex
experiment 1 take 2 results; +plotting switch; some paper updates including references

%!TEX root = main.tex

\section{Experiments and numerical results}

sparse setup + dense setup (in terms of cells, e.g. city center vs rural area)

“sparser”: smallest number of cell towers for which route flow accuracy is pretty good ($<5\%$ error); hopefully somewhat robust to handoff model

“city center”: super dense number of cell towers; handoff model should cause tons of problems here quickly

Motivated by the growth of mobile technology usage and penetration rate, \cite{}