### 2015-02-02 paper/conclusion.tex
minor edits

%!TEX root = main.tex

\section{Conclusion}
Through our investigation of a cell tower handoff noise model, it is clear that handoff dynamics and cellular tower density heavily influence the measured cellpaths and subsequent route flow estimate. Our results indicate that the state estimation method would benefit from an integrated model of handoff dynamics, tailored to the network at hand. Additionally, we observe that handoff dynamics have a magnified effect in denser networks, meaning that state estimation will likely be less accurate. Finally, we conclude that the existing solver, which enforces a single cellpath per route, is too restrictive for the cellpaths introduced by cell tower reassignment. Relaxing this assumption will likely ameliorate the aforementioned effects. Lastly, we plan to perform a full-scale run of the route flow estimation on cellular signaling data from AT\&T in the greater Los Angeles area.